---
title: "Top K predictors"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Top K predictors}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(SimpleEnsembleGroup20)
```
This function utilizes Singular Value Decomposition (SVD) on a predictor matrix to identify and select the top K

K most informative predictors. The selection is based on the magnitudes of the entries in the first right singular vector of the matrix  V


V, which correspond to the predictors. This method is particularly useful in feature selection for high-dimensional data, helping to reduce dimensionality and enhance model interpretability


```{r}
#' @title Select the Top K Informative Predictors Using SVD
#'
#' @description This function performs Singular Value Decomposition (SVD) on the predictor matrix \( X \)
#' and selects the top \( K \) most informative predictors based on the first right singular vector.
#' It checks that \( X \) is either a matrix or a data frame and ensures \( K \) is less than or equal to
#' the number of features in \( X \).
#'
#' @param predictor_matrix A matrix or data frame of predictors.
#' @param top_k The number of top informative predictors to return.
#' @return A list containing a data frame with the top \( K \) predictors, their names, and their scores.
#' @examples
#' \dontrun{
#' data(mtcars)
#' preddata <- mtcars[, -which(names(mtcars) == "mpg")]
#' result <- topk(preddata, 3)
#' print(result$toppreddata)
#' print(result$prednames)
#' print(result$scores)
#' }
#' @export
topk <- function(predictor_matrix, top_k) {
  if (!is.matrix(predictor_matrix) && !is.data.frame(predictor_matrix)) {
    stop("predictor_matrix must be a matrix or a data frame.")
  }

  if (top_k > ncol(predictor_matrix)) {
    stop("top_k must be less than or equal to the number of predictors in predictor_matrix.")
  }

  # Perform Singular Value Decomposition
  svd_result <- svd(predictor_matrix)

  # Extract the right singular vectors (V matrix)
  V_matrix <- svd_result$v

  # Identify the top K informative predictors based on the first right singular vector
  # Considering the absolute values to measure impact regardless of sign
  informative_scores <- abs(V_matrix[, 1])
  top_predictor_indices <- order(informative_scores, decreasing = TRUE)[1:top_k]

  # Retrieve the names of the top predictors
  top_prednames <- colnames(predictor_matrix)[top_predictor_indices]

  # Create a data frame of only the top K predictors
  toppreddata <- predictor_matrix[, top_prednames, drop = FALSE]

  return(list(
    toppreddata = toppreddata,
    prednames = top_prednames,
    scores = informative_scores[top_predictor_indices]
  ))
}


```

Steps Performed:
Data Loading and Preparation: Loads the mtcars dataset and prepares the predictor data by removing the mpg column, which is typically the response variable in regression tasks involving this dataset.
Function Execution: Calls the topk function with the prepared predictor data and a request to identify the top 3 predictors. This function performs SVD to determine which predictors have the most influence based on the first right singular vector.




```{r}
data(mtcars)
preddata <- mtcars[, -which(names(mtcars) == "mpg")]
result <- topk(preddata, 3)
print(result$toppreddata)
print(result$prednames)
print(result$scores)


```


Output Display: Prints three outputs:
result$toppreddata: The subset of the data frame containing only the top 3 predictors.
result$prednames: The names of these top 3 predictors.
result$scores: The scores indicating the relative importance or influence of these predictors based on the SVD.
Expected Outputs:
Top Predictor Data (result$toppreddata): A data frame with only 3 columns, corresponding to the most informative predictors identified by SVD.
Predictor Names (result$prednames): An array of 3 names corresponding to the predictors identified as most informative.
Scores (result$scores): Numeric values representing the scores of these top predictors, indicating the magnitude of their coefficients in the context of the dataset's SVD.
