---
title: "Linear Model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Linear Model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
this is the package
```{r setup}
library(SimpleEnsembleGroup20)
```

This function fits a linear regression model using the least squares method, accommodating both numerical and categorical variables (by converting categorical variables into dummy variables). The model optionally includes an intercept term and offers an optional bagging mechanism. Bagging (Bootstrap Aggregating) can enhance the model's stability and accuracy by averaging results from multiple bootstrap samples, particularly effective in reducing variance and avoiding overfitting in models with high-dimensional data.


```{r}
#' @title Linear Regression Model Fitting with Optional Bagging
#'
#' @description Fits a linear regression model using the least squares method. Supports both numerical
#' and categorical variables by converting categorical variables into dummy variables.
#' When bagging is TRUE, performs bagging to improve the robustness of the model fit.
#'
#' @param y Response variable, numeric vector.
#' @param X Predictor variables, matrix or data frame.
#' @param intercept Logical, if TRUE, an intercept term is added to the model.
#' @param bagging Logical, if TRUE, performs bagging.
#' @param R Integer, number of bootstrap samples for bagging.
#' @return A list containing model details such as coefficients, standard errors,
#'         t-values, p-values, and R-squared values. If bagging is TRUE, returns
#'         a list of aggregated results from the bagging procedure.
#' @importFrom MASS ginv
#' @export
#' @examples
#' data(mtcars)
#' model <- linear(mtcars$mpg, mtcars[, -1], intercept = TRUE, bagging = FALSE)
#' print(model)
linear <- function(y, X, intercept = TRUE, bagging = FALSE, R = 100) {

  if(intercept){
    X <- cbind(Intercept = 1, X)
  }

  if (!bagging) {
    op <- linearinternal(y, X, intercept)
  } else {
    op <- perform_bagging(y, X, function(response_sample, predictor_sample) {
      linearinternal(response_sample, predictor_sample, intercept)
    }, R)
  }

  op$type <- "gaussian"  # Identify the model type for predict_model compatibility
  op$names <- colnames(X)
  return(op)
}


#' Internal function to fit linear model
#'
#' @param y Response variable vector.
#' @param X Design matrix with predictors.
#' @return A list containing model fitting details.

linearinternal <- function(y, X, intercept) {
  if (intercept) {
    X <- cbind(Intercept = 1, X)
  }

  coefficients <- solve(t(X) %*% X) %*% t(X) %*% y
  XtX_inv <- solve(t(X) %*% X)

  # Compute diagnostics
  fitted_values <- X %*% coefficients
  residuals <- y - fitted_values
  rss <- sum(residuals^2)
  tss <- sum((y - mean(y))^2)
  r_squared <- 1 - rss/tss
  adj_r_squared <- 1 - (1 - r_squared) * (length(y) - 1) / (length(y) - ncol(X) - 1)
  se_coefficients <- sqrt(diag(XtX_inv) * rss / (length(y) - ncol(X)))
  t_values <- coefficients / se_coefficients
  p_values <- 2 * pt(-abs(t_values), df = length(y) - ncol(X))
  f_statistic <- (tss - rss) / ncol(X) / (rss / (length(y) - ncol(X) - 1))

  # Summary table construction
  summary_table <- data.frame(
    Estimate = coefficients,
    StdError = se_coefficients,
    tValue = t_values,
    Prt = p_values
  )

  # Set the row names of the summary table to match the coefficient names
  betacols <- colnames(X)
  if (length(coefficients) == length(betacols)) {
    rownames(summary_table) <- betacols
  } else {
    stop("Mismatch in the number of coefficients and predictor names")
  }

  names(coefficients) <- colnames(X)

  # Return a detailed list
  return(list(
    coefficients = coefficients,
    residuals = residuals,
    fitted_values = fitted_values,
    r_squared = r_squared,
    adj_r_squared = adj_r_squared,
    se_coefficients = se_coefficients,
    t_values = t_values,
    p_values = p_values,
    f_statistic = f_statistic,
    df = length(y) - ncol(X),
    rss = rss,
    tss = tss,
    summary = summary_table,
    type = "gaussian"
  ))
}


```

Parameters:
y: Response variable, a numeric vector.
X: Predictor variables, provided as a matrix or data frame.
intercept: Logical, indicates whether to include an intercept term in the model. Defaults to TRUE.
bagging: Logical, specifies whether to perform bagging to enhance model robustness. Defaults to FALSE.
R: Integer, number of bootstrap samples to use for bagging if enabled.
Returns:
A list containing:

coefficients: Model coefficients.
standard errors: Standard errors of the coefficients.
t-values: t-values for the coefficients.
p-values: p-values corresponding to the t-values.
R-squared and Adjusted R-squared: Measures of how well the model fits the data.
F-statistic: A statistic giving an overall significance in the regression.
residuals: The differences between observed and predicted values.
summary: A data frame summarizing the model fit details.
type: A character string indicating the model type for compatibility with predict functions.

```{r}
# Load the mtcars dataset
data(mtcars)

# Fit a linear model using the 'linearinternal' function
model <- linearinternal(mtcars$mpg, as.matrix(mtcars[, -1]), intercept = TRUE)

# Print the model
print(model)

```


